---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loadData, message=FALSE}
require(lubridate)
require(dplyr)
require(lattice)
unzip('activity.zip')
dat <- read.csv('activity.csv')
dat$date <- ymd(dat$date)
```

## What is mean total number of steps taken per day?

```{r stepsPerDay}
stepsPerDay <- summarise(group_by(dat, date), sum(steps))
colnames(stepsPerDay)[2] <- 'steps'
xyplot(steps ~ date, stepsPerDay, type = 'h', main = 'Steps taken per day')
meanSteps <- as.integer(mean(stepsPerDay$steps, na.rm = TRUE))
medianSteps <- median(stepsPerDay$steps, na.rm = TRUE)
```

The mean of the total number of steps taken per day is `r meanSteps`, and the median is `r medianSteps`.

## What is the average daily activity pattern?

```{r stepsPattern}
stepsPattern <- summarise(group_by(dat, interval), mean(steps, na.rm = TRUE))
colnames(stepsPattern)[2] <- 'steps'
xyplot(steps ~ interval, stepsPattern, type = 'l',
       xlab = 'minute of day', ylab = 'steps of 5-minute interval',
       main = 'Average daily activity pattern')
maxInterval <- stepsPattern$interval[which.max(stepsPattern$steps)]
```

The 5-minute interval of `r maxInterval` contains the maximum number of steps on average across all the days in the dataset.

## Imputing missing values

Filling NAs with average daily activity pattern of each interval.
```{r simulateData}
NAsLoc <- which(is.na(dat$steps))
NAsNum <- length(NAsLoc)
dat2 <- dat
for (i in NAsLoc) {
        loc <- which(stepsPattern$interval==dat2$interval[i])
        dat2$steps[i] <- as.integer(stepsPattern$steps[loc])
}
stepsPerDay2 <- summarise(group_by(dat2, date), sum(steps))
colnames(stepsPerDay2)[2] <- 'steps'
xyplot(steps ~ date, stepsPerDay2, type = 'h', 
       main = 'Steps taken per day (with simulated data)')
meanSteps2 <- as.integer(mean(stepsPerDay2$steps))
medianSteps2 <- median(stepsPerDay2$steps)
```

The mean of the total number of steps taken per day in new dataset is `r meanSteps2`, and the median is `r medianSteps2`. After filling NAs with average daily activity pattern of each interval, the mean of the total number of steps taken per day has increased, but the median has decreased.

## Are there differences in activity patterns between weekdays and weekends?

```{r stepsPatternCompare}
dat2$day <- as.factor(weekdays(dat2$date))
levels(dat2$day) <- c(Monday='weekday',
                      Tuesday='weekday',
                      Wednesday='weekday',
                      Thursday='weekday',
                      Friday='weekday',
                      Saturday='weekend',
                      Sunday='weekend')
stepsGroup <- summarise(group_by(dat2, interval, day), mean(steps))
colnames(stepsGroup) <- c('Interval', 'day', 'Number of steps')
xyplot(`Number of steps` ~ Interval | day, stepsGroup,
       layout = c(1, 2), type = 'l')
```
